A Technique for Inter-Catalogue Recommendations

### 2014-06-11 paper2/paper.tex
Minor changes to NeighborhoodRecommender subsection.

\section*{Algorithms}

@@

A key feature of tfidf that is useful here is how terms that appear in many item
description are discounted via the idf factor. Conversely, rare terms will have
larger tfidf values. We now define the $ContentSimilarity$ between items $A$ and
$B$ as
\begin{align}
ContentSimilarity(A,B) = v_A \cdot v_B
\end{align}
where $v_A$ and $v_B$ are the tfidf vector representations of items $A$ and $B$,
respectively.

@@

The $ContentBasedRecommender$ deserves a few more comments. For starters, the
use of an inner product, which is bounded only by the length of the tfidf
vectors, is not fundamental to this algorithm. A bounded similarity function
such as cosine similarity could have been used instead without qualitatively
changing the results discussed later in this report. This is because the lengths
of the product descriptions have little variation, and therefore, the vector
norms of the tfidf vectors have little variation. Finally, we mention that while
this algorithm is quadratic in the number items, it can be executed relatively
efficiently using matrix multiplication, which trades time for space (i.e.
memory). Furthermore, confining our experiments to individual categories reduces
the impact of the quadratic time complexity. In our case, Macy's entire
catalogue consists of over $60,000$ products, making a quadratic time algorithm
challenging. However, the average category has about two thousand products,
which is a significantly more manageable.

\subsection*{Leveraging Recommendations}
This $ContentBased Recommender$ is limited in a few ways. For one, it attempts
to capture similarity between products by looking exclusively at the words in
their descriptions. One issue here is that brands tend to use different 
vocabularies to describe the their products. One may refer to a color as 
``off-white'' while another calls the same color ``ivory.'' To the $ContentBased
Recommender$, these words are as different as ``black'' and ``white.'' Another 
issue is that the best recommendation for a product may not be the most similar
product. For instance, a striped black-and-white button-up shirt could be 
an excellent recommendation for a blue checkered button-up, even though the two 
have ostensibly different descriptions.

We can alleviate some of these issues by leveraging the recommendations we
already have between products within the same catalogue. One approach is to 
consider the {\em neighborhood} of each product, that is a product along 
with each of its neighbors in the recommendation graph. With this, we create a
new similarity function called $NeighborhoodSimilarity$. This function uses the 
$ContentSimilarity$ function from before to compare every pair of products 
between the two neighborhoods:

\begin{align}
NeighborhoodSimilarity(P_1, P_2) &= \sum\limits_{S\in
Neigh(P_1)}\sum\limits_{T\in Neigh(P_2)}
ContentSimilarity(S, T) 
\end{align}
where $Neigh(P)$ is the neighborhood of Product P. We use this similarity
function to construct the $Neighborhood Recommender$, which is analgous to the
$ContentBased Recommender$.

Assuming that the edges of the recommendation graphs are in fact good
recommendations, this approach can help both of the issues brought up with the
$ContentBased Recommender$. Since a neighborhood has many products, each 
describing themselves using potentially different vocabularies, there is a 
higher likelihood that the neighborhood contains multiple synonyms for the
underlying concept that describes their similarity. This decreases the
likelihood of false negatives due to different vocabularies. Furthermore,
the recommendation graphs capture some of the user behavior that cannot be
expressed in product descriptions. Re-examining the button-up shirts example, 
the neighborhood of a striped shirt may include checkered shirts and vise-versa.
In this case, the two neighborhoods would share the words ``striped'' and 
``checkered'' and thus the two previously dissimilar shirts would now correctly
be deemed similar.

\subsection*{Latent Concepts}
Although the $Neighborhood Recommender$ alleviates some of the issues of 
content-based similarity, it is still primarily based on the descriptions of
products. Ideally, we would like to recommend products that are similar in more
fundamental ways. If a dress is elegant and modern, we would like to recommend 
another dress that is elegant and modern, even if those words are not explicitly
in the descriptions of either. These types of underlying concepts can be 
intangible and difficult to define. However, if we assume that the given
recommendations are truly based on such latent concepts, we can attempt to learn 
them by analyzing the graphs.

@@

In all, the latent-concept approach consists of four phases. First, we simulate user
sessions by randomly walking the recommendation graphs. Second, we learn concept
spaces for products in each catalogue via SVD. Third, we find a mapping between
concept-spaces using the aggregated descriptions of all products. And lastly, we
find the nearest neighbors of each product in the concept-space of the other 
catalogue concept-space. These nearest neighbors become the predicted
recommendations between products across the two catalogues. We call this method
the $LatentConcept Recommender$

>>>>

\section*{Recommender Algorithms}

@@

A key feature of tfidf that is useful here is how terms that appear in many item
descriptions are discounted via the idf factor. Conversely, rare terms will have
larger tfidf values. We now define the $ContentSimilarity$ between items $A$ and
$B$ as
\begin{align}
ContentSimilarity(A,B) = v_A \cdot v_B
\end{align}
where $v_A$ and $v_B$ are the tfidf vector representations of items $A$ and $B$,
respectively.

@@

The $ContentBasedRecommender$ deserves a few more comments. For starters, the
use of an inner product, which is bounded only by the length of the tfidf
vectors, is not fundamental to this algorithm. A bounded similarity function
such as cosine similarity could have been used instead without qualitatively
changing the results discussed later in this report. This is because the lengths
of the product descriptions have little variation, and therefore, the vector
norms of the tfidf vectors have little variation. Finally, we mention that while
this algorithm is quadratic in the number items, it can be executed relatively
efficiently using matrix multiplication, which trades time for space (i.e.
memory). Furthermore, confining our experiments to individual categories reduces
the impact of the quadratic time complexity. In our case, Macy's entire
catalogue consists of over $60,000$ products, making a quadratic time algorithm
challenging. However, the average category has about two thousand products,
which is significantly more manageable.

\subsection*{\em NeighborhoodRecommender}
The $ContentBasedRecommender$ is limited in that it attempts to capture
similarity between products by looking exclusively at the terms in their
descriptions. One issue here is that brands tend to use different vocabularies
to describe their products. One may refer to a color as ``off-white'' while
another calls the same color ``ivory.'' To the $ContentBasedRecommender$ these
terms are as different as ``black'' and ``white.'' Another issue is that the
best recommendation for a product may not even be the most similar product. For
instance, a striped black-and-white button-up shirt could be an excellent
recommendation associated with a blue checkered button-up, even though the two
products are clearly different and would presumably use signficantly different
terms their descriptions.

We can begin to address these issues by leveraging the recommendations we
already have between products within a single partition. One approach is to
consider the {\em neighborhood} of each product, that is a product along with
each of its immediate neighbors in the recommendation graph. With this, we
create a new similarity function called $NeighborhoodSimilarity$. This function
uses the $ContentSimilarity$ function from before to compare every pair of
products between the two neighborhoods of a pair of products $A$ and $B$:
\begin{align}
NeighborhoodSimilarity(A, B) = \sum\limits_{S\in
Neigh(A)}\sum\limits_{T\in Neigh(B)}
ContentSimilarity(S, T) 
\end{align}
where $Neigh(A)$ is the set of items that share edges with item $A$. We use this
similarity function to construct the $NeighborhoodRecommender$, which is
an extension of the $ContentBasedRecommender$. The algorithm is the same as
before except now we choose a recommendation associated with items $A$ and $B$
by maximnizing
\begin{align}
ContentSimilarity(A,B) + \eta \cdot NeighborhoodSimilarity(A, B),
\end{align}
where $\eta$ is a parameter. We find that $\eta = 0.1$ yields good results for
this algorithm.

Assuming that the edges of the recommendation graphs are in fact good
recommendations, this approach can help both of the shortcomings of
$ContentBasedRecommender$ previously mentioned. Since a neighborhood includes
several items, each with descriptions that potentially use different
vocabularies, there is a higher likelihood that the neighborhood contains
multiple synonyms for the underlying concept that captures their similarity. At
the same time, this decreases the likelihood of false negatives due to different
vocabularies. Presumably, a retailer's recommendation graph include information
from user behavior that is not expressed in product descriptions. Re-examining
the button-up shirts example, the neighborhood of a striped shirt may include
checkered shirts and vice versa.  In this case, the two neighborhoods might
share the words ``striped'' and ``checkered'' and thus the two previously
dissimilar shirts would now be considered more similar.

\subsection*{\em LatentConceptRecommender}
Although the $NeighborhoodRecommender$ alleviates some of the issues of 
content-based similarity, it is still primarily based on the descriptions of
products. Ideally, we would like to recommend products that are similar in more
fundamental ways. If a dress is elegant and modern, we would like to recommend 
another dress that is elegant and modern, even if those words are not explicitly
in the descriptions of either. These types of underlying concepts can be 
intangible and difficult to define. However, if we assume that the given
recommendations are truly based on such latent concepts, we can attempt to learn 
them by analyzing the graphs.

@@

In all, the latent-concept approach consists of four phases. First, we simulate
user sessions by randomly walking the recommendation graphs. Second, we learn
concept spaces for products in each catalogue via SVD. Third, we find a mapping
between concept-spaces using the aggregated descriptions of all products. And
lastly, we find the nearest neighbors of each product in the concept-space of
the other catalogue concept-space. These nearest neighbors become the predicted
recommendations between products across the two catalogues. We call this method
the $LatentConceptRecommender$
